---
title: "EC48T_Project"
author: "Atakan"
date: "20/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About

This is our group project (3 contributing members) for our "Financial Modelling and Its Applications Using R" course. 

This report's aim is trying to find a cross correlation between RUB/TRY and WTI/TRY price through evaluating "Adjusted" prices.

We collected data, which is from 2010 and 2021, from yahoo finance, divided them into 4 time frames, and evaluating each one of them.

Also, this report presents a basic financial model based on the trading opportunity that may come out with WTI crude oil prices and the currency. 

Through this model, we wanted to know whether we can make profit out of this opportunity.

After the calculations, we found a positive correlation and trading opportunity. However, we make very small profit out of it. 

We used "quantmod" and "tidyverse" packages in our project.

```{r, message=FALSE, warning=FALSE}
library(quantmod)
library(tidyverse)
```

## Collecting the data

  We are collecting the data from yahoo finance. However, yahoo finance does not have RUB/TRY and WTI(according to TRY) data so we basically took RUB/USD, TRY/USD, and WTI (in terms of USD). 
  
  Then, we adjusted those to have RUB/TRY and WTI values in TRY through assigning new variables.

```{r, message=FALSE, warning=FALSE}
rub <- getSymbols("RUB=X", src = "yahoo", from = as.Date("2010-01-01"), to = as.Date("2021-12-31"), auto.assign = F)
try <- getSymbols("TRY=X", src = "yahoo", from = as.Date("2010-01-01"), to = as.Date("2021-12-31"), auto.assign = F)
wti_crude_oil <- getSymbols("CL=F", src = "yahoo", from = as.Date("2010-01-01"), to = as.Date("2021-12-31"), auto.assign = F)

rubtry <- try/rub

wti <- wti_crude_oil*try
```

## Cleaning the data

  To clean the "rubtry" and "wti" data, we need to create dataframes out of them.

```{r}
rubtry <- data.frame(Date = index(rubtry), rubtry, row.names = NULL)

wti <- data.frame(Date = index(wti), wti, row.names = NULL)
```
  
  First, we cleaned the rubtry data.
  
  We decided that we do not need the "Volume" of rubtry, we omitted the null values, and assigned much more cleaner column names.
  
  We found only one outlier in the "Adjusted" column which is a incorrect as we controlled it, so we removed it from our dataframe.

```{r}
rubtry <- select(rubtry, -TRY.X.Volume)

rubtry <- na.omit(rubtry)

colnames(rubtry) <- c("Date", "Open", "High", "Low", "Close", "Adjusted")

if (max(rubtry["Adjusted"]) > 3) {
  maxrow <- which.max(rubtry$Adjusted)
  rubtry <- rubtry[c(-(maxrow)), ]
}
```
  
  For the wti data, we did not encounter with outliers or incorrect data. 
  
  We ommitted the null values and assigned new column names.
  
```{r}
wti <- na.omit(wti)

colnames(wti) <- c("Date","Open", "High", "Low", "Close", "Volume", "Adjusted")
```
  
  We found out that row numbers of both dataframe are not matched with each other since we ommitted null values and an outlier (Also, the dataset we get may be problematic.)
  
  To solve this problem, we match each row on a dataframe with another dataframe based on their "Date" data so that if one date consists of both rubtry and wti data, then it will not be removed.
  
```{r}
xrubtry <- data.frame()

for (i in 1:length(rubtry$Date)){
  
  if (any(rubtry$Date[i] == wti$Date)){
    xrubtry <- rbind(xrubtry, rubtry[i, ])
  }
}

rubtry <- xrubtry

xwti <- data.frame()

for (i in 1:length(wti$Date)){
  
  if (any(wti$Date[i] == rubtry$Date)){
    xwti <- rbind(xwti, wti[i, ])
  }
}

wti <- xwti
```
  
  Now, the row numbers are matched and we can calculate the cross correlation.

## Filtering the data

  For our research, we took the data from 2010 to 2021 and we divided it to 4 time frame to investigate each one of them (2010-2012, 2013-2015, 2016-2018, 2019-2021). 
  
```{r}
rubtry_1 <- filter(rubtry, between(Date, as.Date("2010-01-01"), as.Date("2012-12-31")))
rubtry_2 <- filter(rubtry, between(Date, as.Date("2013-01-01"), as.Date("2015-12-31")))
rubtry_3 <- filter(rubtry, between(Date, as.Date("2016-01-01"), as.Date("2018-12-31")))
rubtry_4 <- filter(rubtry, between(Date, as.Date("2019-01-01"), as.Date("2021-12-31")))

wti_1 <- filter(wti, between(Date, as.Date("2010-01-01"), as.Date("2012-12-31")))
wti_2 <- filter(wti, between(Date, as.Date("2013-01-01"), as.Date("2015-12-31")))
wti_3 <- filter(wti, between(Date, as.Date("2016-01-01"), as.Date("2018-12-31")))
wti_4 <- filter(wti, between(Date, as.Date("2019-01-01"), as.Date("2021-12-31")))
```

## Plots

  Here are the plots of each time frame that we filtered.

```{r, echo=FALSE}
coeff_1 <- 2500
coeff_2 <- 2800
coeff_3 <- 2500
coeff_4 <- 3000

plot_1 <- ggplot() +
  geom_line(data = rubtry_1, aes(x = Date, y = Adjusted, colour = "RUB/TRY")) +
  geom_line(data = wti_1, aes(x = Date, y = Adjusted / coeff_1, colour = "WTI")) +
  ggtitle("2010-2012 RUB/TRY and WTI Prices") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_y_continuous(
    name = "RUB/TRY",
    sec.axis = sec_axis(~.*coeff_1, name = "WTI")
  ) +
  scale_colour_manual("", 
                      breaks = c("RUB/TRY", "WTI"),
                      values = c("black", "red"))

plot_2 <- ggplot() +
  geom_line(data = rubtry_2, aes(x = Date, y = Adjusted, colour = "RUB/TRY")) +
  geom_line(data = wti_2, aes(x = Date, y = Adjusted / coeff_2, colour = "WTI")) +
  ggtitle("2013-2015 RUB/TRY and WTI Prices") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_y_continuous(
    name = "RUB/TRY",
    sec.axis = sec_axis(~.*coeff_2, name = "WTI")
  ) +
  scale_colour_manual("", 
                      breaks = c("RUB/TRY", "WTI"),
                      values = c("black", "red"))

plot_3 <- ggplot() +
  geom_line(data = rubtry_3, aes(x = Date, y = Adjusted, colour = "RUB/TRY")) +
  geom_line(data = wti_3, aes(x = Date, y = Adjusted / coeff_3, colour = "WTI")) +
  ggtitle("2016-2018 RUB/TRY and WTI Prices") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_y_continuous(
    name = "RUB/TRY",
    sec.axis = sec_axis(~.*coeff_3, name = "WTI")
  ) +
  scale_colour_manual("", 
                      breaks = c("RUB/TRY", "WTI"),
                      values = c("black", "red"))

plot_4 <- ggplot() +
  geom_line(data = rubtry_4, aes(x = Date, y = Adjusted, colour = "RUB/TRY")) +
  geom_line(data = wti_4, aes(x = Date, y = Adjusted / coeff_4, colour = "WTI")) +
  ggtitle("2019-2021 RUB/TRY and WTI Prices") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_y_continuous(
    name = "RUB/TRY",
    sec.axis = sec_axis(~.*coeff_4, name = "WTI")
  ) +
  scale_colour_manual("", 
                      breaks = c("RUB/TRY", "WTI"),
                      values = c("black", "red"))

print(plot_1)
print(plot_2)
print(plot_3)
print(plot_4)

```
  
  We may say that each plot seems RUB/TRY and WTI/TRY are highly correlated, but this claim should be supported by calculations.
  
## Analysis

  We calculated the cross correlation and make linear regression to each time frame.

### Cross Correlation

```{r}
ccf_1 <- print(ccf(rubtry_1["Adjusted"], wti_1["Adjusted"]))
ccf_2 <- print(ccf(rubtry_2["Adjusted"], wti_2["Adjusted"]))
ccf_3 <- print(ccf(rubtry_3["Adjusted"], wti_3["Adjusted"]))
ccf_4 <- print(ccf(rubtry_4["Adjusted"], wti_4["Adjusted"]))
```

  With respect to these calculations, we can see that 2016-2018 time frame has the highest correlation by "0.942". Then the 2013-2015 comes as the second, 2019-2012 comes as the third, and the last one is 2010-2012.
  
  Clearly, we can supported the claim that RUB/TRY and WTI/TRY are positively and highly correlated within each time frame, therefore they are correlated from 2010 to 2021.
  
  Within these findings, we can now create a financial model that can be evaluating this correlation, finding a trading opportunity, and making profit out of it. 

  But before making the model, we need to do the linear regression approach.
  
### Linear Regression

```{r}
test_1 <- lm(rubtry_1$Adjusted~wti_1$Adjusted)
test_2 <- lm(rubtry_2$Adjusted~wti_2$Adjusted)
test_3 <- lm(rubtry_3$Adjusted~wti_3$Adjusted)
test_4 <- lm(rubtry_4$Adjusted~wti_4$Adjusted)

print(test_1)
print(test_2)
print(test_3)
print(test_4)
```

## Financial Model

  We created a financial model that calculates the profit if there is a trading opportunity.
  
  The model's code is as follows:
  
```{r}
ProfitCalculate <- function(wti_price, rub_price, coeff, int, day_lag, initial_budget, trade_amount)
{
  if(length(wti_price)!=length(rub_price))
  {
    return("Prices series should have the same length")
  }
  
  rub_hold <- 0
  wti_hold <- 0
  budget <- initial_budget
  
  start <- day_lag+1
  end <- length(rub_price)-1
  
  for(i in start:end)
  {
    
    if(rub_price[i] >= wti_price[i-day_lag])
    {
      budget + 100
      rub_hold <- rub_hold - rub_price/100
    }
    
    else if(rub_price[i] < wti_price[i-day_lag])
    {
      budget - 100
      rub_hold <- rub_hold + rub_price[i]/100
    }
  }
  
  calced_profit <- budget + rub_price[length(rub_price)] * rub_hold - initial_budget
  
  return(paste("Profit = ", calced_profit))
}
```
  
  In this function, we considered our wti and rubtry prices, coefficients and intercepts of the time frame that we calculate and we assumed a day lag as we kind of applying backtesting approach, initial budget, and amount of trade in our research.
  
  Also, we wanted to remove our last time frame which is 2019-2021 in this report because its coefficients are a really odd comparing with other time frames.

```{r}
profit_1 <- ProfitCalculate(wti_price = wti_1$Adjusted, rub_price = rubtry_1$Adjusted,
                            coeff = test_1$coefficients[2], int = test_1$coefficients[1],
                            day_lag = 10, initial_budget = 1000, trade_amount = 100)

profit_2 <- ProfitCalculate(wti_price = wti_2$Adjusted, rub_price = rubtry_2$Adjusted,
                            coeff = test_2$coefficients[2], int = test_2$coefficients[1],
                            day_lag = 10, initial_budget = 1000, trade_amount = 100)

profit_3 <- ProfitCalculate(wti_price = wti_3$Adjusted, rub_price = rubtry_3$Adjusted,
                            coeff = test_3$coefficients[2], int = test_3$coefficients[1],
                            day_lag = 10, initial_budget = 1000, trade_amount = 100)

print(profit_1)
print(profit_2)
print(profit_3)
```

  We found that we can make profit, but it is very small, nearly 2-3%.
